\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-05 k-planar.tex
added more commens on 1-planar section

\note{DW}{We can motivate \cref{1-planar} since treewidth 3 is best possible, and the planarity of $H$ is essential in the proof of \cref{1PlanarQueue}. Shall we argue this in the response? Or do we remove \cref{sec-1-planar}? We could keep the statement of \cref{1-planar} in the paper and refer to the arXiv paper for the proof. What do you think? All this depends on whether we can fix the errors below. }

Let $G$ be an edge-maximal 1-plane multigraph with no two parallel edges on the boundary of a single face.  Here, edge-maximal should be taken to mean that, if any two vertices $v$ and $w$ appear on a common face\footnote{The \defin{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.  We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.  We assume that no two edges incident to a common vertex cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}

@@

\referee{2}{Page 13. Perhaps I am misunderstanding something, but I do not see why
'none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of
$G$.'  For example, see the attached PDF for a picture where $vw$, $xw$, $wy$,
and $yv$ are all crossed by other edges of $G$.\\
\includegraphics[width=60mm]{kite}}
\note{DW}{We need to explain more about how edge-maximality fixes this problem. In this case we can draw other $xv$, $vy$, $yw$ $wx$ edges close to the central crossing edges.}
\note{DW}{The graph drawing literature defines a 1-planar graph to be \defin{optimal} if it is obtained from a plane quadrangulation by adding the two crossing edges across each face. Optimal 1-planar graphs have the maximum number of edges out of all $n$-vertex 1-planar graphs.	But I am pretty sure there are edge-maximal 1-planar graphs that are not optimal. We should check that our proof is okay on these examples. }
Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.

>>>>

\note{DW}{We can motivate \cref{1-planar} since treewidth 3 is best possible, and the planarity of $H$ is essential in the proof of \cref{1PlanarQueue}. Futher motivation comes from the open problem at the end of the paper. Shall we argue this in the response? Or do we remove \cref{sec-1-planar}? We could keep the statement of \cref{1-planar} in the paper and refer to the arXiv paper for the proof. What do you think? All this depends on whether we can fix the errors below. I vote to fix the errors, add more motivation for \cref{1-planar}, keep the theorem and the proof. }

Let $G$ be an edge-maximal 1-plane multigraph with no two parallel edges on the boundary of a single face.  Here, edge-maximal should be taken to mean that, if any two vertices $v$ and $w$ appear on a common face\footnote{The \defin{faces} of an embedded graph $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$.  We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.} $F$, then there is an edge $vw\in E(G)$ that is contained in the boundary of $F$.  \note{DW}{Each face of $G$ is one of two types, either bounded by three edges (none of which are crossed), or bounded by one edge that is not crossed plus portions of two crossing edges. Right? If this is true, we should say it here, which would help to clear up the confusion below. Add a figure.} 
We assume that no two edges incident to a common vertex cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}

\note{DW}{I suggest we say that it follows from Euler's formula that $G$ is finite. (To see this, say $G$ has $n$ vertices, $m$ non-crossed edges, and $k$ crossings. Let $G'$ be the plane graph obtained from $G$ by adding a dummy vertex at each crossing. So $G'$ is a plane multigraph in which every face is bounded by a 3-cycle. It follows from Euler's formula that $m+4k = |E(G')| = 3( |V(G')|-2) = 3 ( n+k-2)$, implying	$m+k = 3 (n-2)$ and $|E(G)| = m+2k \leq 6(n-2)$.) }

\note{DW}{It seems to be that $G$ can be obtained from a plane multigraph with faces of size 3 or 4 such that no edge is in two 3-faces, by adding a pair of crossing edges across each 4-face. I think this characterises edge-maximal 1-planar multigraphs with no 2-faces. Right? Does this viewpoint help? Is it known?}

\note{DW}{Say $G$ is obtained from a plane multigraph by making each face a clique. Is $G\subseteq H \boxtimes P \boxtimes K_{f(\omega(G)}$? for some treewidth 3 graph $H$.}

@@

\referee{2}{Page 13. Perhaps I am misunderstanding something, but I do not see why
'none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of
$G$.'  For example, see the attached PDF for a picture where $vw$, $xw$, $wy$,
and $yv$ are all crossed by other edges of $G$.\\
\includegraphics[width=60mm]{kite}}
\note{DW}{We need to explain more about how edge-maximality fixes this problem. In this case we can draw other $xv$, $vy$, $yw$ $wx$ edges close to the central crossing edges.} Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
